### 2017-12-07 paper/sexSelectionSpacing-ver1.tex
Edit of model requirements/set-up

Requirements for method and the intuition behind each:  
 - Account for censoring (hazard model)
 - Work both with and without sex selection (competing risk part)
 - Allow for changes in use of sex selection within spell (flexible specification of baseline hazard ?)
 - Capture that the shape of the hazard function differ across groups
   (non-proportionality); needs a better description, maybe using example
   of shorter spacing and/or use of sex selection. Differences in parity
   progression likelihood.

I divide a woman's reproductive life into one or more spells.
Each spell covers the period from one birth, or marriage for the first spell,
to the following birth; I define how the duration is measured in more 
detail in Section \ref{sec:data}.
A basic problem is that some spells' durations are censored, either
because the survey took place before the end birth for the spell or because
the couple has finished childbearing.
The standard approach, which I follow here, is to use a hazard models to 
address the censoring of spells.

Prior research on birth spacing have used hazard models where the exit 
state was the birth of a child. 
This works when there is no sex selection---making the sex of the child 
a random event---but not here because at least some couples use 
sex selection. 
To capture that the sex of children are no longer a random event, I instead
use a competing risk framework with two possible exit states: either a boy 
or a girl is born.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

Proportional hazard models, where covariate have an multiplicative 
effect on the hazard rate, are more efficient than non-proportional 
models, but only provided that the proportionality assumption holds.
The problem here is that a proportional hazard model does not allow covariates
to have different sized effects at different times within a spell and therefore
cannot capture differences in the shape of the hazard functions between 
different groups, and if the proportionality assumption does not hold 
the result is a potentially substantial bias in estimates.
It is unlikely---even in the absence of prenatal sex 
determination---that the effect of, for example, the sex composition
of prior births have the same effect throughout the entire spell.
[son preference and parity progression also leads to differences!]
Furthermore, any bias from the proportionality assumption is likely 
exacerbated by the introduction of prenatal sex determination.
First, sex-selective abortions affects birth spacing and its use
vary across groups.
Second, the use of sex selection may vary within a spell, which means 
that the effect of covariate may vary within the spell as well.
Finally, the use of a non-proportional specification, together with a 
flexible baseline hazard, also mitigates any potential effects of 
unobserved heterogeneity \citep{Dolton1995}.
My proposed model is a discrete time, non-proportional competing risk 
hazard model with two exit states: either a boy or a girl is born.

>>>>

% Requirements for method and the intuition behind each:  
%  - Account for censoring (hazard model)
%  - Work both with and without sex selection (competing risk part)
%  - Allow for changes in use of sex selection within spell (flexible specification of baseline hazard ?)
%  - Capture that the shape of the hazard function differ across groups
%    (non-proportionality); needs a better description, maybe using example
%    of shorter spacing and/or use of sex selection. Differences in parity
%    progression likelihood.

I divide women's reproductive lives into spells, which are the periods from 
one birth, or marriage for the first spell, to the following birth
(I define how spell duration is measured in Section \ref{sec:data}).
A basic problem is that some spells' duration are censored, either
because the survey took place before the birth that would end the spell 
or because the couple has finished childbearing.
The standard approach, which I follow here, is to use a hazard models to 
address the censoring of spells.

Prior research on birth spacing have used hazard models where the only 
exit state was the birth of a child. 
This works when there is no sex selection---making the sex of the child 
a random event---but not here because at least some couples use 
sex selection. 
To capture that the sex of children are no longer necessarily a random event, 
I instead use a competing risk framework with two possible exit states: 
either a boy or a girl is born.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

Proportional hazard models, where covariates have an multiplicative 
effect on the hazard rate, are more efficient than non-proportional 
models, but only provided that the proportionality assumption holds.
Because proportional hazard models do not allow covariates to have 
different sized effects at different times within a spell,
they cannot capture differences in the shape of the hazard functions 
between different groups, and if the proportionality assumption does 
not hold the estimates may be biased.
It is unlikely---even in the absence of prenatal sex determination---that 
the effect of, for example, the sex composition of prior births have 
the same effect throughout the entire spell.
This is especially the case for later spells where different
sex composition of prior births can lead to substantial difference 
both in the likelihood of progressing to the next birth and
how soon couples want their next child if they are going to have one.
Any bias from the proportionality assumption is likely 
exacerbated by the introduction of prenatal sex determination.
First, sex-selective abortions affect birth spacing and use
vary across groups.
Second, the use of sex selection may vary within a spell, which means 
that the effects of covariates vary within the spell as well.
Because of these concerns I use a non-proportional specification.
The use of a non-proportional specification also has the benefit 
that it mitigates any potential effects of unobserved heterogeneity 
when used in conjunction with a flexible baseline hazard \citep{Dolton1995}.
My proposed model is a discrete time, non-proportional, competing risk 
hazard model with two exit states: either a boy or a girl is born.
